\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-11-21 paper/main.tex
WIP fixes to section 3

Across different versions of a given package, functions may be
changed, depreciated, added, or renamed.  In addition to changing the
behaviors of active computations, saved objects created using one
version of a package may be incompatible with another version of the
same package.  For example, the popular \texttt{pandas}~\cite{McKi10}
package prior to version 0.25.0 included the \texttt{Panel} datatype
for storing 3-dimensional arrays.  Since version 0.20.0,
however, the \texttt{Panel} datatype has been depreciated, and the
datatype was removed entirely in version 0.25.0.  If one had saved a dataset in a
\texttt{Panel} object (created using an older version of
\texttt{pandas}), e.g., by serializing the \texttt{Panel} and saving
it to disk using \texttt{pickle}, that dataset could not be read by
any version of \texttt{pandas} from 0.25.0 or beyond.  These
incompatibilities are not limited solely to traditional forms of data.
For example, saved model states and other objects may reference
functions, attributes, classes, and other objects that are not present
across all versions of their associated package.

@@

The example provided in Figure~\ref{fig:illustrative-example}
demonstrates how the \texttt{davos} package can be used to circumvent
these incompatibilities by carefully controlling which versions of
each package are used in different parts of the notebook.  The example
shows how data and a model saved using since-depreciated components of
the \texttt{pandas} and \texttt{scikit-learn}~\cite{PedrEtal11}
packages may be loaded in (using older versions of each package) and
manipulated or analyzed (using more recent versions of each package).

After installing and importing \texttt{davos} (lines 1--2), we first smuggle two
utilities for interacting with local files in the code below. The
\texttt{smuggle} statement in line 4 loads the \texttt{is\_file()}
function from the Python standard library's \texttt{os.path}
module. Standard library modules are included with all Python
distributions, so this line is functionally equivalent to an
\texttt{import} statement and does not need or benefit from an onion
comment. Line 5 loads the \texttt{joblib} package~\cite{Varo10},
installing it first, if necessary. Since \texttt{joblib}'s I/O
interface has historically remained stable and backwards-compatible
across releases, requiring that users have a particular exact version
installed would likely be unnecessarily restrictive. However, a
\textit{future} release might introduce some breaking change.  The
onion comment in line 5 helps ensure the analysis notebook continues
to run properly in the future, by limiting allowable versions to those
already released when the code was written:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example1}
\end{center}
Line 7 then uses the \texttt{davos.config} object to enable
\texttt{davos}'s \texttt{auto\_rerun} option before smuggling the next
two packages: \texttt{NumPy}~\cite{HarrEtal20} and
\texttt{pandas}. Because these packages rely heavily on custom C data
types, loading the particular versions from the onion comments may
require restarting the notebook kernel if different versions had been previously
imported during the same interpreter session (see
Sec.~\ref{subsec:config}).
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example2}
\end{center}
Setting the \texttt{auto\_rerun} attribute to \texttt{True} also
benefits the (potential) installation of \texttt{pandas} on the next
lines:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example3}
\end{center}
If we suppose that the data contained in \texttt{data-old.pkl} is
stored in a \texttt{Panel} object (i.e., created in a version of
\texttt{pandas} prior to 0.25.0), then we would not be able to load in
that object with newer versions of \texttt{pandas}.  Line 11 ensures
that an older version of \texttt{pandas} will be imported, enabling
the data to be read in (and, in line 13, saved out to a \texttt{.csv}
file, which is compatible with newer versions of \texttt{pandas}).

>>>>

Across different versions of a given package, particular modules, functions, and other objects 
may be updated, removed, renamed, or otherwise altered in some manner.  In addition to changing the
behaviors of active computations, these changes can render saved objects created using one
version of a package incompatible with other versions of the
same package.  For example, the popular \texttt{pandas}~\cite{McKi10}
library used to include the \texttt{Panel} data structure for storing 3-dimensional arrays.  
Since version 0.20.0, however, the \texttt{Panel} class has been deprecated, and in 
version 0.25.0, it was removed entirely. If a user had a dataset stored in a \texttt{Panel} object
(created using an older version of \texttt{pandas}) and saved it to their disk (e.g., for later reuse
or to share with other users) by serializing the \texttt{Panel} with Python's \texttt{pickle} protocol, 
that dataset could not be read by any version of \texttt{pandas} from 0.25.0 or beyond. These
incompatibilities are also not limited solely to traditional forms of data.
For example, saved model states and other objects may reference
modules, functions, attributes, classes, or other objects that may not be identical (or even present)
across all versions of their associated package.

@@

The example provided in Figure~\ref{fig:illustrative-example}
demonstrates how the \texttt{davos} package can be used to circumvent
these incompatibilities by carefully controlling which versions of
each package are used in different parts of the notebook.  The example
shows how a dataset and model that require now-incompatible components of
the \texttt{pandas} and \texttt{scikit-learn}~\cite{PedrEtal11}
packages may be loaded in (using older versions of each package) and used
alongside more recent versions of each package that provide new and improved functionality.
When included at the top of a Jupyter notebook, the code in Figure~\ref{fig:illustrative-example}
ensures that these objects will be loaded successfully and analyzed using the same set
of package versions, no matter when or by whom the notebook is run.

After installing and importing \texttt{davos} (lines 1--2), we first smuggle two
utilities for interacting with local files in the code below. The
\texttt{smuggle} statement in line 4 loads the \texttt{is\_file()}
function from the Python standard library's \texttt{os.path}
module. Standard library modules are included with all Python
distributions, so this line is functionally equivalent to an
\texttt{import} statement and does not need or benefit from an onion
comment. Line 5 loads the \texttt{joblib} package~\cite{Varo10},
installing it first, if necessary. Since \texttt{joblib}'s I/O
interface has historically remained stable and backwards-compatible
across releases, requiring that users have a particular exact version
installed would likely be unnecessarily restrictive. However, a
\textit{future} release might introduce some breaking change.  The
onion comment in line 5 helps ensure the analysis notebook continues
to run properly in the future by limiting allowable versions to those
already released when the code was written:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example1}
\end{center}
Line 7 then uses the \texttt{davos.config} object to enable
\texttt{davos}'s \texttt{auto\_rerun} option before smuggling the next
two packages: \texttt{NumPy}~\cite{HarrEtal20} and
\texttt{pandas}. Because these packages rely heavily on custom C data
types, loading the particular versions from the onion comments may
require restarting the notebook kernel if different versions had been previously
imported during the same interpreter session (see
Sec.~\ref{subsec:config}).
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example2}
\end{center}
Setting the \texttt{auto\_rerun} attribute to \texttt{True} is particularly useful
for the installation of \texttt{pandas} in the next
lines:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example3}
\end{center}
If we suppose that the data contained in \texttt{data-old.pkl} is
stored in a \texttt{Panel} object, then we must use a version of 
\texttt{pandas} prior to 0.25.0 (i.e., the version in which the \texttt{Panel} 
class was removed) to be able to load it in. Line 11 ensures
that an older version of \texttt{pandas} will be imported, enabling
the data to be read in (and, in line 13, saved out to a \texttt{.csv}
file, which is compatible with newer versions of \texttt{pandas}).
